---
title: "Reproducible Research - Course Project 1"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading and preprocessing the data
```{r loading}
Act <- read.csv("activity.csv")
```

### What is mean total number of steps taken per day?  
1. Calculate the total number of steps taken per day
```{r total steps per day}
ActStep <- tapply(Act$steps, Act$date, sum, na.rm=TRUE)
ActStep
```

2. Make a histogram of the total number of steps taken each day
```{r step per day hist}
hist(ActStep, breaks=20, xlab="Steps", main="Total number of steps taken per day")

```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r mean_median per day}
mean(ActStep)
median(ActStep)
```

### What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r daily pattern plot}
ActPat <- aggregate(steps ~ interval, data=Act, mean, na.rm=TRUE)
plot(ActPat, type = "l", xlab="Interval (min)", ylab="Steps", main=" Average number of steps")

```

  2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max step}
ActPat[which.max(ActPat$steps),]
```

### Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r counting NAs}
sum(is.na(Act))
```

  2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.  

- Replacing NA to mean of 5-minute interval

  3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r imputing NA}
mergedAct <- merge(Act, ActPat, by="interval")
mergedAct$steps.x[is.na(mergedAct$steps.x)] = mergedAct$steps.y[which(is.na(mergedAct$steps.x))]
sum(is.na(mergedAct))
```

  4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r imputing NA hist}
mergedActStep <- tapply(mergedAct$steps.x, mergedAct$date, sum)
hist(mergedActStep, breaks=20, xlab="Steps", main="Total number of steps taken per day")
```

```{r imputing NA mean median}
mean(mergedActStep)
median(mergedActStep)
```

  - Both histograms have similar pattern except decreased 0 values in the dataset imputing NA compared to the original dataset in which NA set as 0 by default. Mean of total number of steps per day increased by replacing those 0s to each mean of 5-minute interval.
  
  
### Are there differences in activity patterns between weekdays and weekends?
  1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r weekday}
mergedAct$date <- as.Date(mergedAct$date, "%Y-%m-%d")
mergedAct$wkday <- weekdays(mergedAct$date)
wd <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
mergedAct$wkday <- factor(mergedAct$wkday %in% wd, levels=c(FALSE, TRUE), labels=c("weekend", "weekday"))
str(mergedAct$wkday)
```

  2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r weekday plot}
stepWd <- aggregate(steps.x ~ interval+wkday, mergedAct, mean)
library(lattice)
xyplot(steps.x ~ interval | wkday, data=stepWd, layout=c(1, 2), type="l", ylab="Steps", xlab="Interval (min)")
```




